A fast and simple primal-dual algorithm for computing Nash equilibria and best response strategies in two-person zero-sum sequential games with imcomplete information

### 2014-11-29 onepage.tex
ehn

\begin{abstract}
% LESS THAN 200 WORDS !!!!!
% 
% While medical imaging typically provides massive amounts of
% data, the extraction of relevant information in a given
% applicative context remains a difficult challenge.
% XXX : don't call it SPM but just "predictive regions or predictive
% maps"
In this manuscript we present a simple primal-dual algorithm for computing Nash equilibria and best reponse strategies in  sequential games with
imcomplete information and perfect recall. The proposed algorithm derives from the primal-dual scheme of A. Chambolle and T. Pock, applied to the
equilibrium and best response strategy problems for the sequence form representation of the game. We use a dualization trick to render all
proximity operators computable in closed-form.

@@

\begin{IEEEkeywords}
  game theory; Nash equilibrium; two-person sequential game; imcomplete information; perfect recall; sequence-form representaion; best response strategy;
  convex-optimization; dualization; primal-dual algorithms
\end{IEEEkeywords}

\section{Introduction}
%\label{sec:intro}
\subsection{Notation and Terminology}
Give positive integers $m$ and $n$, $\mathbb{R}^{n}$ denotes the $n$-dimensional real vector space, and $\mathbb{R}^{m \times n}$
the space of all $m$-by-$n$ real matrices. $0_{m,n}$ denotes the $m$-by-$n$ matrix of zeros.
$\mathbb{R}^n_+$ := $\{x \in \mathbb{R}^n|x_j \ge 0 \text{ }  \forall j\}$ is the $n$-dimensional \textit{nonnegati
ve orthant}.
For a vector $x \in \mathbb{R}^n$, $\|x\|$ denotes the $2$-\textit{norm} of $x$ defined by $\|x\| := \sqrt{x^Tx}$.
$(x)_+$ denotes its point-wise maximum with 0. Note that $(x)_+ \in \mathbb{R}^n_+$.
For example, $((-2, \pi))_+ = (max(-2, 0), max(\pi, 0)) = (0, \pi)$. The \textit{spectral norm} of a matrix $K$, denoted $\|K\|$, is defined by
$\|K\| := \underset{x \ne 0}{max}\text{ }\frac{\|Kx\|}{\|x\|}$. Equivalently, $\|K\|$ is the largest \textit{singular value} of $K$, i.e the largest
\textit{eigen-value} of $K^TK$. Given a convex subset $C$ of $\mathbb{R}^n$, $i_C$ denotes its \textit{indicator function} defined by
$i_C(x) = 0$ if $x \in C$ and $i_C(x) = +\infty$ otherwise. Finally, given a \textit{proper} convex function (i.e a convex function which takes a finite
value at least a one point) $f : \mathbb{R}^n \rightarrow [0, +\infty]$, and a positive real number $\tau$, the \textit{proximal operator} of $f$ of rank $\tau$,
denoted $\text{prox}_{\tau f}$ is the function which maps a point $x \in \mathbb{R}^n$ to the (necessarily unique) solution of the problem
\begin{equation}
  \underset{z \in \mathbb{R}^n}{argmin}\text{ }\frac{1}{2}\|z - x\|^2 + \tau f(z)
\end{equation}

@@

\subsection{Statement of the problem}
We are interested in the problem of computing a Nash equilibrium for a two-person zero-sum sequential game with imcomplete information
and perfect recall (e.g Texas Hold'em Poker, etc.). In the \textit{sequence-form representation} of the game, the problem has the following saddle-point formulation
\begin{equation}
  \underset{y \in Q_2}{minimize}\text{ }\underset{x \in Q_1}{maximize}\text{ }{y^TAx}
  \label{eq:opt_pb}
\end{equation}

where $Q_j := \{z \in \mathbb{R}_+^{n_j}|E_jz=e_j\}$, for $j \in \{1, 2\}$. As usual, the ``minimize-maximize'' notation in problem
\eqref{eq:opt_pb} means that a pair $(x^*, y^*) \in Q_1 \times Q_2$ is a solution iff
\begin{equation}
  {y^*}^TAx \le y^TAx \le y^TAx^*, \forall (x, y) \in Q_1 \times Q_2
\end{equation}
Such pairs correspond to the Nash equilibria of the game.

We recall that $E_j$ is a $p_j$-by-$n_j$ matrix whose
entries are $-1$, $0$, or $+1$, and $e_j := (1, 0, 0, ..., 0)$ is a vector of length $p_j$. We also recall that $E_1$ and $e_1$ (resp. $E_2$ and $e_2$)
encode linear constraints on our (resp. the opponent's)  ``admissible'' realization $x$ (resp. $y$). $A$ is the payoff matrix, of size $n_2 \times n_1$.

@@

\section{The proposed algorithm}
We preprose to solve problem \eqref{eq:opt_pb} using the primal-dual scheme of A. Chambolle and T. Pock.
Though this scheme has recently gained considerable popularity in the signal processing community, to the best of our knowledge,
this is the first time it is being applied to game theoretic problems.

@@

Now, observe that
\begin{eqnarray*}
  i_{Q_j}(z) = i_{\mathbb{R}^{n_j}_+}(z) + \underset{w \in \mathbb{R}^{n_j}}{max}\text{}{w^T(E_jz - e_j)}, \forall z \in \mathbb{R}^{n_j},
\end{eqnarray*}
and so problem \eqref{eq:opt_pb} can be re-written as

\begin{equation}
  \underset{y \in \mathbb{R}^{n_2}, v\in \mathbb{R}^{p_1}}{minimize}\text{ }\underset{x \in \mathbb{R}^{n_1}, u \in \mathbb{R}^{p_2}}{maximize}
           {\begin{bmatrix}y\\v\end{bmatrix}^TK\begin{bmatrix}x\\u\end{bmatrix} + G(y, v) - F(x, u)}
  \label{eq:my_opt_pb}
\end{equation}

where:
\begin{equation}
  \left .
  \begin{split}
    K :=
    \left[
      \begin{array}{c|c}
        A & E_1^T \\ \hline
        E_2 & 0_{p_2, p_1}
      \end{array}
      \right] \in \mathbb{R}^{(n_2 + p_1) \times (n_1 + p_2)} \\
    %%\begin{bmatrix}A \text{ } E_1^T\\ E_2 \text{ } 0\end{bmatrix} \in \mathbb{R}^{(n_2 + p_1) \times (n_1 + p_2)}\\
      G: \mathbb{R}^{n_2} \times \mathbb{R}^{p_1} \rightarrow [0, +\infty], (y, v) \mapsto i_{\mathbb{R}^{n_2}_+}(y) - v^Te_1\\
      F: \mathbb{R}^{n_1} \times \mathbb{R}^{p_2} \rightarrow [0, +\infty], (x, u) \mapsto i_{\mathbb{R}^{n_1}_+}(x) - u^Te_2
  \end{split}
  \right\}
\end{equation}

>>>>

\begin{abstract}
% LESS THAN 200 WORDS !!!!!
% 
% While medical imaging typically provides massive amounts of
% data, the extraction of relevant information in a given
% applicative context remains a difficult challenge.
% XXX : don't call it SPM but just "predictive regions or predictive
% maps"
In this manuscript, we present a simple primal-dual algorithm for computing Nash equilibria and best reponse strategies in  sequential games with
imcomplete information and perfect recall. The proposed algorithm derives from the primal-dual scheme of A. Chambolle and T. Pock, applied to the
equilibrium and best response strategy problems in the sequence-form representation of the game. Our technique is to use a dualization
trick and unconstrain the saddle-point problem. All proximal operators in the resulting primal-dual algorithm are simple,
only involving nonnegative thresholding and translations.
The convergence and convergence rate of our algorithm derive explicitly from the general Chambolle-Pock primal-dual scheme.

@@

\begin{IEEEkeywords}
  game theory; Nash equilibrium; two-person zero-sum game; sequential game; imcomplete information; perfect recall; sequence-form representaion;
  best response strategy; convex-optimization; dualization; proximal operator; primal-dual algorithms
\end{IEEEkeywords}

\section{Introduction}
%\label{sec:intro}

@@

\subsection{Statement of the problem}
We are interested in the problem of computing a Nash equilibrium for a two-person zero-sum sequential game with imcomplete information
and perfect recall (e.g Texas Hold'em Poker, etc.). In the \textit{sequence-form representation} of the game, the problem has the following saddle-point formulation
\begin{equation}
  \underset{y \in Q_2}{minimize}\text{ }\underset{x \in Q_1}{maximize}\text{ }{x^TAy}
  \label{eq:opt_pb}
\end{equation}

where $Q_j := \{z \in \mathbb{R}^{n_j}|z \ge 0, E_jz=e_j\}$, for $j \in \{1, 2\}$. As usual, the ``minimize-maximize'' notation in problem
\eqref{eq:opt_pb} means that a pair $(x^*, y^*) \in Q_1 \times Q_2$ is a solution iff
\begin{equation}
  {y^*}^TAx \le y^TAx \le y^TAx^*, \forall (x, y) \in Q_1 \times Q_2
\end{equation}
Such pairs correspond to the Nash equilibria of the game.

We recall that $E_j$ is a $p_j$-by-$n_j$ matrix whose
entries are $-1$, $0$, or $+1$, and $e_j := (1, 0, 0, ..., 0)$ is a vector of length $p_j$. We also recall that $E_1$ and $e_1$ (resp. $E_2$ and $e_2$)
encode linear constraints on our (resp. the opponent's)  ``admissible'' realization $x$ (resp. $y$). $A$ is the payoff matrix, of size $n_1 \times n_2$.

@@

\section{The proposed algorithm}
We preprose to solve problem \eqref{eq:opt_pb} using the primal-dual scheme of A. Chambolle and T. Pock.
Though this scheme has recently gained considerable popularity in the signal processing community, to the best of our knowledge,
this is the first time it is being applied to compute Nash equilibria.

We first introduce some fundamental notations, and then proceed to derive the algorithm. The convergence and convergence rate
of our algorithm derive explicitly from the general Chambolle-Pock primal-dual scheme.

\subsection{Notation and Terminology}
Give positive integers $m$ and $n$, $\mathbb{R}^{m \times n}$ denotes
the space of all $m$-by-$n$ real matrices. $0_{m,n}$ denotes the $m$-by-$n$ matrix of zeros.
$\mathbb{R}^n_+$ := $\{x \in \mathbb{R}^n|x_j \ge 0 \text{ }  \forall j\}$ is the $n$-dimensional \textit{nonnegative orthant}.
For a vector $x \in \mathbb{R}^n$, $\|x\|$ denotes the $2$-\textit{norm} of $x$ defined by $\|x\| := \sqrt{x^Tx}$.
$(x)_+$ denotes its point-wise maximum with 0. Note that $(x)_+ \in \mathbb{R}^n_+$.
For example, $((-2, \pi))_+ = (max(-2, 0), max(\pi, 0)) = (0, \pi)$. The \textit{spectral norm} of a matrix $K$, denoted $\|K\|$, is defined by
\footnote{Equivalently, $\|K\|$ is the largest \textit{singular value} of $K$, i.e the largest \textit{eigen-value} of $K^TK$.}
$\|K\| := \underset{x \ne 0}{max}\text{ }\frac{\|Kx\|}{\|x\|}$.

Given a \textit{convex subset} $C$ of $\mathbb{R}^n$, $i_C$ denotes its \textit{indicator function} defined by
$i_C(x) = 0$ if $x \in C$ and $i_C(x) = +\infty$ otherwise. Note that $i_{C \cap D} = i_C + i_D$. Finally, given a \textit{proper convex
function} (i.e a convex function which takes a finite
value at least a one point) $f : \mathbb{R}^n \rightarrow [0, +\infty]$, and a positive real number $\tau$, the \textit{proximal operator} of $f$ of rank $\tau$,
denoted $\text{prox}_{\tau f}$ is the function which maps a point $x \in \mathbb{R}^n$ to the (necessarily unique) solution of the problem
\begin{equation}
  \underset{z \in \mathbb{R}^n}{argmin}\text{ }\frac{1}{2}\|z - x\|^2 + \tau f(z)
\end{equation}

@@

\subsection{Derivation of the algorithm}
Observe that $i_{Q_j}(z) = i_{\mathbb{R}^{n_j}_+}(z) + \underset{w \in \mathbb{R}^{n_j}}{max}\text{}{w^T(E_jz - e_j)}, \forall z \in \mathbb{R}^{n_j}$,
and so problem \eqref{eq:opt_pb} can be re-written as

\begin{equation}
  \underset{y \in \mathbb{R}^{n_2}, v\in \mathbb{R}^{p_1}}{minimize}\text{ }\underset{x \in \mathbb{R}^{n_1}, u \in \mathbb{R}^{p_2}}{maximize}
           {\begin{bmatrix}x\\u\end{bmatrix}^TK\begin{bmatrix}y\\v\end{bmatrix} + G(y, v) - F(x, u)}
  \label{eq:my_opt_pb}
\end{equation}

where:
\begin{equation}
  \left .
  \begin{split}
    K :=
    \left[
      \begin{array}{c|c}
        A & E_1^T \\ \hline
        E_2 & 0_{p_2, p_1}
      \end{array}
      \right] \in \mathbb{R}^{(n_1 + p_2) \times (n_2 + p_1)} \\
    %%\begin{bmatrix}A \text{ } E_1^T\\ E_2 \text{ } 0\end{bmatrix} \in \mathbb{R}^{(n_2 + p_1) \times (n_1 + p_2)}\\
      G: \mathbb{R}^{n_2} \times \mathbb{R}^{p_1} \rightarrow [0, +\infty], (y, v) \mapsto i_{\mathbb{R}^{n_2}_+}(y) - v^Te_1\\
      F: \mathbb{R}^{n_1} \times \mathbb{R}^{p_2} \rightarrow [0, +\infty], (x, u) \mapsto i_{\mathbb{R}^{n_1}_+}(x) - u^Te_2
  \end{split}
  \right\}
\end{equation}
